---
title: "OECD Data Collection and Exploration"
goal: get the OECD Data Explorer API to work and pull relevant fishing and mining data
---

### [0] prep

```{r setup, echo=FALSE}
if (!require(pacman)) install.packages('pacman')
library(pacman)
p_load(tidyverse, RColorBrewer, ggplot2, OECD)
```

### [1] data

```{r data finding}
# if ds does not exist, then get it
if (!exists("ds")) {
  ds <- get_datasets()
}

ds_fish <- search_dataset("Fisheries", ds) #comment
ds_mine <- search_dataset("Mine", ds)
ds_gdp <- search_dataset("GDP", ds)


ds_fish_codes <- ds_fish[,1]
ds_mine_codes <- ds_mine[,1]
ds_gdp_codes <- ds_gdp[,1]


fish_ind <- get_dataset("FISH_INDICATORS")
fish_rnd <- get_dataset("FISH_PAT_RD")
fish_trade <- get_dataset("FISH_TRADE")
fish_empl <- get_dataset("FISH_EMPL")
fish_patdev <- get_dataset("FISH_PAT_DEV")

```

```{r fish data merging}

# join the five datasets above into one using column merge by country and time
fish_data <- merge(fish_ind, fish_rnd, by.x = c("COUNTRY", "Time"), by.y = c("COU", "Time"))
fish_data <- merge(fish_data, fish_trade, by = c("COUNTRY", "Time"))
fish_data <- merge(fish_data, fish_empl)

```

```{r mining data}

## get data
mine_data <- get_dataset(ds_mine_codes)
```
```{r gdp data}
## need to look through and find which of the datasets we want to pull
#lets use get_dataset() to grab each of the datasets in ds_gdp_codes one by one and show a tibble
# print(ds_gdp[,2], n=28)
# gdp_data1 <- get_dataset(ds_gdp_codes[[1]][1])
# head(gdp_data1)
# gdp_data2 <- get_dataset(ds_gdp_codes[[1]][2])
# head(gdp_data2)
# gdp_data3 <- get_dataset(ds_gdp_codes[[1]][3])
# head(gdp_data3)
# gdp_data4 <- get_dataset(ds_gdp_codes[[1]][4])
# head(gdp_data4)

## given a link, access OECD API and save the table as a dataframe
link <- "https://sdmx.oecd.org/public/rest/data/OECD.SDD.NAD,DSD_NAMAIN10@DF_TABLE1_EXPENDITURE,1.0/A...........?dimensionAtObservation=AllDimensions"

# https://sdmx.oecd.org/public/dataflow/DF_TABLE1_EXPENDITURE/1.0/
# https://sdmx.oecd.org/public/dataflow/DF_TABLE1_EXPENDITURE/1.0/A.AUS.......XDC.V..?startPeriod=2019&dimensionAtObservation=AllDimensions
df <- readsdmx::read_sdmx((link))
class(df) <- c("tbl_df", "tbl", "data.frame")
df

na_data <- get_dataset("DSD_NAMAIN10")

```



```{r gdp data cleaning}

```

### [2] plots

### 2.1 Time Series Analysis of GDP

This section focuses on visualizing the time series of GDP contributions and growth rates. These plots will help us understand the trends and fluctuations in GDP over time.

```{r time-series-plots, echo=TRUE, message=FALSE, warning=FALSE}

## relative change rates
ggplot(df_gdp_calcs_1,
       aes(x = Year, y = GDP_change_relative,
           color=Country)) +
  geom_line()+
  geom_point(size=.7,alpha=0.6)+
  facet_wrap(~Industry_simple,scales="free_y")+
  labs(title = paste0("Rates of Change in Relative GDP Contributions"),
       x = NULL,
       y = "Annual Change (%)",
       color="Country") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1,size=10),
        axis.text.y = element_text(size=10),
        axis.title = element_text(size=14),
        legend.text = element_text(size=12),
        legend.title = element_text(size=12),
        plot.title = element_text(size=15),
        legend.position = "none")
# ggsave("plots/line_relchange.jpg")

## gdp contributions
ggplot(df_gdp_calcs_1,
       aes(x = Year, y = GDP_perc_norm,
           color=Country)) +
  geom_line()+
  geom_point(size=.7,alpha=0.6)+
  facet_wrap(~Industry_simple,scales="free_y")+
  labs(title = paste0("Relative GDP Contributions"),
       x = NULL,
       y = "% GDP",
       color="Country") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1,size=10),
        axis.text.y = element_text(size=10),
        axis.title = element_text(size=14),
        legend.text = element_text(size=12),
        legend.title = element_text(size=12),
        plot.title = element_text(size=15))
# ggsave("plots/line_relgdp.jpg")

```

### 2.2 boxplots of GDP and changes

```{r boxplots, echo=TRUE, message=FALSE, warning=FALSE}
## relative change rates
ggplot(df_gdp_calcs_1,
       aes(x = Country_short, y = GDP_change_relative,color=Industry_simple)) +
  geom_abline(slope = 0, intercept = mean(df_gdp_calcs$GDP_change_relative[df_gdp_calcs$Industry_simple == "Fishing"], na.rm = TRUE),
              color = brewer.pal(5,"Dark2")[1], linetype = "dashed") + 
  geom_abline(slope = 0, intercept = mean(df_gdp_calcs$GDP_change_relative[df_gdp_calcs$Industry_simple == "Mining"], na.rm = TRUE),
              color = brewer.pal(5,"Dark2")[2], linetype = "dashed") + 
  geom_boxplot(alpha=0.8)+
  geom_point(size=.7,alpha=0.6)+
  scale_color_brewer(palette="Dark2")+
  # facet_wrap(~Industry_simple,scales="free_y")+
  labs(title = paste0("Rates of Change in Relative GDP Contributions"),
       x = NULL,
       y = "Annual Change (%)",
       color="Industry") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1,size=10),
        axis.text.y = element_text(size=10),
        axis.title = element_text(size=14),
        legend.text = element_text(size=12),
        legend.title = element_text(size=12),
        plot.title = element_text(size=15))
# ggsave("plots/box_relchange.jpg")

##
ggplot(df_gdp_calcs_1,
       aes(x = Country_short, y = GDP_perc_norm,color=Industry_simple)) +
  geom_abline(slope = 0, intercept = mean(df_gdp_calcs$GDP_perc_norm[df_gdp_calcs$Industry_simple == "Fishing"], na.rm = TRUE),
              color = brewer.pal(5,"Dark2")[1], linetype = "dashed") + 
  geom_abline(slope = 0, intercept = mean(df_gdp_calcs$GDP_perc_norm[df_gdp_calcs$Industry_simple == "Mining"], na.rm = TRUE),
              color = brewer.pal(5,"Dark2")[2], linetype = "dashed") + 
  geom_boxplot()+
  geom_point(size=.7,alpha=0.6)+
  # facet_wrap(~Industry_simple, scales = "free_y") +
  labs(title = paste0("Relative GDP Contributions"),
       x = NULL,
       y = "% GDP",
       color="Industry") +
  scale_color_brewer(palette="Dark2")+
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1,size=10),
        axis.text.y = element_text(size=10),
        axis.title = element_text(size=14),
        legend.text = element_text(size=12),
        legend.title = element_text(size=12),
        plot.title = element_text(size=15))
# ggsave("plots/box_relgdp.jpg")


```

### 2.3 stacked bar chart GDP contributions - aggregate

```{r bar-plots, echo=TRUE, message=FALSE, warning=FALSE}
# prep data - remove any annual timeseries variables
gdp_contr <- group_by(df_gdp_calcs_1,Country) %>% 
  mutate(both_rel_mean=sum(unique(GDP_perc_norm_mean))) %>% ungroup() %>%
  group_by(Country,Industry_simple,Aggregate_level) %>%
  mutate(industry_total_gva=sum(GVA_total)) %>% ungroup() %>%
  select(-c("Industry","Year","GVA_total","Base_year","GDP_change_abs_mean","GDP_change_rel_mean", "Effective_year_GDP","GDP_annual","GDP_perc_norm","GDP_change_absolute","GDP_change_relative")) %>%
  distinct() %>%
  arrange(desc(both_rel_mean)) %>%
  mutate(across(where(is.character), ~factor(., levels = unique(.)))) %>%
  group_by(Country, Industry_simple) %>%
  mutate(percent_weight = industry_total_gva / sum(industry_total_gva[Aggregate_level == 1])) %>% ungroup() %>%
  mutate(agg1_yn = ifelse(Aggregate_level==1,as.character(1),as.character(0)))

# percentage (%)
ggplot(gdp_contr,
       aes(x = Country_short, y = GDP_perc_norm_mean, fill=Industry_simple)) +
  geom_bar(stat = "identity") +
  labs(title = "Economic Contribution of Fishing and Mining Sectors",
       x = NULL,
       y = "% GDP",
       fill = NULL) +
  scale_fill_brewer(palette = "Set2") + 
  scale_y_continuous(limits=c(0,50),labels=seq(0,50,5),
                     breaks=seq(0,50,5),expand=c(0,0))+
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1,size=12),
        axis.text.y = element_text(size=10),
        axis.title = element_text(size=12),
        legend.text = element_text(size=12),
        plot.title = element_text(size=15),
        legend.position = c(.85,.85), legend.background = element_rect(fill="white",size=0.2))
# ggsave(paste0("plots/plot_barchart_percgdp_",gdp_type, ".jpg"))
```

### 2.3 stacked bar chart GDP contributions - individual

```{r bar-plots, echo=TRUE, message=FALSE, warning=FALSE}
## plot 2 individual sectors
# process data
gdp_contr_fish <- filter(gdp_contr,Industry_simple=="Fishing") %>%
  arrange(desc(GDP_perc_norm_mean)) %>%   mutate(Country = factor(Country, levels = unique(Country)))
fish_range<- range(gdp_contr_fish$GDP_perc_norm_mean)
gdp_contr_mine <- filter(gdp_contr,Industry_simple=="Mining") %>%
  arrange(desc(GDP_perc_norm_mean)) %>%   mutate(Country = factor(Country, levels = unique(Country)))
mine_range<- range(gdp_contr_mine$GDP_perc_norm_mean)

# plot
p_fish <- ggplot(gdp_contr_fish, aes(x = reorder(Country_short, -GDP_perc_norm_mean), y = GDP_perc_norm_mean, fill = Industry_simple)) +
  geom_bar(stat = "identity", show.legend = FALSE,fill=brewer.pal(8, "Set2")[1]) +
  labs(title = paste0(unique(gdp_contr_fish$Industry_simple)),
       x = NULL,y = "% GDP") +
  scale_y_continuous(limits = c(floor(fish_range[1] / 5) * 5,ceiling(fish_range[2] / 5) * 5),expand=c(0,0))+
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 7),
        axis.text.y = element_text(size = 8),
        axis.title = element_text(size = 10),
        legend.text = element_text(size = 10),
        legend.title = element_text(size = 10),
        plot.title = element_text(size = 10),
        strip.text = element_text(size = 10, face = "bold"))
p_mine <- ggplot(gdp_contr_mine, aes(x = reorder(Country_short, -GDP_perc_norm_mean), y = GDP_perc_norm_mean, fill = Industry_simple)) +
  geom_bar(stat = "identity", show.legend = FALSE,fill=brewer.pal(8, "Set2")[2]) +
  labs(title = paste0(unique(gdp_contr_mine$Industry_simple)),
       x = NULL, y = NULL) +
  scale_y_continuous(limits = c(floor(mine_range[1] / 5) * 5,ceiling(mine_range[2] / 5) * 5),expand=c(0,0))+
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 7),
        axis.text.y = element_text(size = 8),
        axis.title = element_text(size = 10),
        legend.text = element_text(size = 10),
        legend.title = element_text(size = 10),
        plot.title = element_text(size = 10),
        strip.text = element_text(size = 10, face = "bold"))

cowplot::plot_grid(p_fish,p_mine) # combine em
# ggsave("plots/plot_barchart_percgdp_separate.jpg")

```

### 2.4 scatterplots fishing vs. mining contributions

```{r scatter-plots, echo=TRUE, message=FALSE, warning=FALSE}
fishing_data <- filter(df_gdp_calcs_1, Industry_simple == "Fishing") %>%
  rename(Fishing_Contribution = GDP_perc_norm_mean,
         Fishing_Rel = GDP_change_rel_mean) 
mining_data <- filter(df_gdp_calcs_1, Industry_simple == "Mining") %>%
  rename(Mining_Contribution = GDP_perc_norm_mean,
         Mining_Rel = GDP_change_rel_mean) 

# Perform the left join, grab relevant columns, omit na's
combined_data <- full_join(fishing_data, mining_data, by = c("Country","Year")) %>%
  select(Country, Country_iso.y,Fishing_Contribution, Mining_Contribution, Fishing_Rel, Mining_Rel) %>%
  distinct() %>% rename(Country_iso=Country_iso.y) %>%
  filter(rowSums(is.na(.)) == 0)

# run clustering (4_makeclusters.R) to add membership
data_scaled_contr <- scale(combined_data[,c("Fishing_Contribution","Mining_Contribution")])
set.seed(123)  # Setting a random seed for reproducibility
k <- 4
kmeans_result_contr <- kmeans(data_scaled_contr, centers = k, nstart = 25)
# add cluster assignments
combined_data$cluster_contr <- as.factor(kmeans_result_contr$cluster)

## plot ##
# plot mean gdp %
scat_gdp_mean <- ggplot(data = combined_data, 
                        aes(x = Fishing_Contribution, y = Mining_Contribution,
                            color=cluster_contr)) + # for kmeans clusering
  geom_abline(slope = 1, intercept = 0, color = "lightgrey", linetype = "dashed") + # Add a line with slope of 1
  geom_point() +
  geom_text(aes(label = Country_iso), size=4,
            nudge_x = 0, nudge_y = .8, check_overlap = FALSE) + # Add labels
  scale_x_continuous(limits = c(0,30),labels=c(seq(0,30,5)),breaks=c(seq(0,30,5)))+
  scale_y_continuous(limits = c(0,30),labels=c(seq(0,30,5)),breaks=c(seq(0,30,5)))+
  scale_color_brewer(palette = "Dark2") + 
  theme_minimal() + 
  theme(legend.position="none",
        axis.text = element_text(size=14),
        axis.title = element_text(size=12),
        plot.title = element_text(size=12)) +
  labs(x = "Fishing", y = "Mining",title="Contribution to GDP (%)",
       subtitle=paste0(k," k-means clusters"))
plot(scat_gdp_mean)
# ggsave(paste0("plots/scatter_gdp_",gdp_type,".jpg"))
```

```{r scatter-plots, echo=TRUE, message=FALSE,warning=FALSE}
## plot mean relative growth
data_scaled_rel <- scale(combined_data[,c("Fishing_Rel","Mining_Rel")])
set.seed(123)  # Setting a random seed for reproducibility
k <-4
kmeans_result_rel <- kmeans(data_scaled_rel, centers = k, nstart = 25)
# add cluster assignments
combined_data$cluster_rel <- as.factor(kmeans_result_rel$cluster)

mine_range_rel <- range(combined_data$Mining_Rel) # make ranges
fish_range_rel <- range(combined_data$Fishing_Rel) 

scat_gdp_rel <- ggplot(data = combined_data, 
                       aes(x = Fishing_Rel, y = Mining_Rel,
                           color=cluster_rel)) +
  geom_hline(yintercept=0, color = "gray60") + 
  geom_vline(xintercept=0, color = "gray60") + 
  geom_point() + 
  geom_text(aes(label = Country_iso), size=4,
            nudge_x = 0, nudge_y = 2, check_overlap = FALSE) + # Add labels
  scale_x_continuous(limits = c(floor(fish_range_rel[1] / 5) * 5,ceiling(fish_range_rel[2] / 10) * 10))+
  scale_y_continuous(limits = c(floor(mine_range_rel[1] / 5) * 5,ceiling(mine_range_rel[2] / 10) * 10))+
  scale_color_brewer(palette = "Dark2") + 
  theme_minimal() + 
  theme(legend.position="none",
        axis.text = element_text(size=10),
        axis.title = element_text(size=14),
        plot.title = element_text(size=15)) +
  labs(x = "Fishing", y = "Mining",title="Average Annual Relative GDP Change (%)",
       subtitle=paste0(k," k-means clusters"))
plot(scat_gdp_rel)
# ggsave(paste0("plots/scatter_relchange_",gdp_type,".jpg"))
```

### 2.5 check data completion by aggregate level ~facet_wrap()

``` {r barcharts-2level, echo=TRUE, message=FALSE, warning=FALSE}
ggplot(df_gdp_calcs, aes(x = Country_short, y = GDP_perc_norm_mean,
                            fill=Industry_simple),
       show.legend=FALSE) +
  geom_bar(stat = "identity",position="dodge") +
  scale_fill_brewer(palette = "Dark2") + 
  labs(title = "Data completeness: aggregate levels",
       x = "Country",
       y = "% GDP",
       fill = "Industry") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
                legend.position = c(0.85,0.3)) +
  facet_wrap(~Aggregate_level,scales="free")
# ggsave("plots/barchart_aggregates.jpg")

```
